\sf Retraction-based first-order feasible methods for difference-of-convex programs with smooth inequality and simple geometric constraints

### 2022-04-23 SQP_Retract_rev1.tex
Minor edits

\section{Introduction}
We consider the following difference-of-convex (DC) optimization problem with smooth inequality and geometric constraints:
\begin{equation}\label{P0}
  \begin{array}{rl}
\min\limits_{x\in\R^n} & P(x) := P_1(x) - P_2(x)\\
{\rm s.t.} & g_i(x)\leq 0,~~ i = 1,\ldots, m,\\
           & x\in C,
  \end{array}
\end{equation}
where $P_1:\R^n\rightarrow\R$ and $P_2:\R^n\rightarrow\R$ are continuous convex functions, $C\subseteq \R^n$ is a nonempty compact convex set, $g_i:\R^n\rightarrow\R$ is smooth with Lipschitz gradient on $C$ for each $i$, and the feasible set $\mathcal{F} :=\{x\in C:\; g_i(x)\le 0, ~ i = 1,\ldots, m\}$ is nonempty. This model problem covers many important optimization problems arising from diverse areas. For example, one can model the compressed sensing (CS) problem with Cauchy noise by minimizing the difference of $\ell_1$-norm and $\ell_2$-norm \cite{YiLH15} or its variants subject to an inequality constraint defined by the so-called Lorentzian norm \cite{CaBA10}, and the set $C$ in \eqref{P0} can be used for modeling other prior information such as bounds or nonnegativity; more discussions on this particular application can be found in Section~\ref{sec6.2} below. Problem~\eqref{P0} is typically challenging to solve because it is in general nonconvex and nonsmooth.% optimization problem.

@@

More explicitly, we consider the following special case of \eqref{P0}:
\begin{equation}\label{P1}
  \begin{array}{rl}
\min\limits_{x\in\R^n} & P(x) := P_1(x) - P_2(x)\\
{\rm s.t.} & \ell(Ax - b)\leq \sigma,\\
           & x\in C,
  \end{array}
\end{equation}
where $P_1$ and $P_2$ are convex and continuous functions, $C$ is a nonempty compact convex set, $A\in\R^{p\times n}$, $b\in\R^p$,  $\sigma >0$, the set ${\cal F}_1:= \{x\in C:\; \ell(Ax - b)\le \sigma\}\neq \emptyset$, and $\ell$ also satisfies the following Assumption~\ref{gassum}:
\begin{assumption}\label{gassum}
The function $\ell:\R^p\to \R$ is defined as
\[
\ell(u):= \sum_{i=1}^{p} {\color{blue} \varphi}(u_i^2),
\]
where ${\color{blue} \varphi}:\R_+\rightarrow \R_+$ satisfies the following conditions:
\begin{enumerate}[{\rm (i)}]
  \item ${\color{blue} \varphi}$ is a continuous concave function with ${\color{blue} \varphi}(0) = 0$;
  \item ${\color{blue} \varphi}$ is continuously differentiable with positive derivative on $(0, \infty)$;
  \item The limit $\vartheta:= \lim\limits_{t\downarrow 0}{\color{blue} \varphi}'(t)$ exists, and the function $t\mapsto {\color{blue} \varphi}'_+(t)$ is locally Lipschitz on $[0,\infty)$.\footnote{Here, and throughout this paper, we use ${\color{blue} \varphi}'_+(t)$ to denote the right hand derivative of ${\color{blue} \varphi}$ at $t$, which is defined as ${\color{blue} \varphi}'_+(t)=\lim\limits_{h \rightarrow 0^+}\frac{{\color{blue} \varphi}(t+h)-{\color{blue} \varphi}(t)}{h}$.}
\end{enumerate}
\end{assumption}

@@

{\color{blue}We compare Algorithm~\ref{SQP_retract2} and Algorithm~\ref{ESQM} on solving random instances of \eqref{E4}. Specifically, we generate two matrices ${\bf A}_{\rm re}\in \R^{p\times n}$ and ${\bf A}_{\rm im}\in \R^{p\times n}$ with i.i.d standard Gaussian entries, let
\begin{equation*}
A := \left[ \begin{matrix}
{\bf A}_{\rm re} & -{\bf A}_{\rm im}\\
{\bf A}_{\rm im} & {\bf A}_{\rm re}
\end{matrix} \right]
\end{equation*}
and then normalize each column of $A$. Next, we generate two vectors ${\bf u}\in\R^k$ and ${\bf v}\in\R^k$ with i.i.d standard Gaussian entries. We choose an index set $I$ of size $k$ at random and define the original sparse signal ${\bf z}_{\rm orig}\in {\mathbb{C}}^{n}$ by setting ${\bf z}_j := {\bf u}_j + i {\bf v}_j$ for all $j \in I$ and ${\bf z}_j = 0$ otherwise. We then let
\[
\xorig = \begin{bmatrix}
  {\bf z}_{\rm re}\\
  {\bf z}_{\rm im}
\end{bmatrix}\ \ {\rm and}\ \ b = A\xorig + 0.005\hat{\epsilon},
\]
where ${\bf z}_{\rm re}$ and ${\bf z}_{\rm im}$ are respectively the real and imaginary parts of ${\bf z}$, and $\hat{\epsilon}$ is a random vector with each entry independently following the Cauchy$(0, 1)$ distribution, i.e., $\hat{\epsilon}_{j} = \tan(\pi(\widetilde{\epsilon}_j - \frac{1}{2}))$ with each $\widetilde{\epsilon}_j\in\R^{2p}$ being uniformly chosen in $[0, 1]$. Finally, we set $\overline{\sigma} = 1.2\|0.005\hat{\epsilon}\|_{LL_2,\gamma}$ with $\gamma = 0.05$.}

@@

In our numerical experiments, we let $\mu =0.95$, $j = 2$, $\gamma = 0.05$, {\color{blue} $(p,n,k) = (360i,1280i,60i)$ }with $i\in \{2, 4, 6, 8, 10\}$. For each $i$, we generate 20 random instances as described above. The computational results averaged over the {\color{blue} $20$ random instances} are shown in Table~\ref{table2}. In the table, Algorithm~\ref{SQP_retract2} is denoted by FPA$_{\rm retract}$, Algorithm~\ref{ESQM} with different $\delta$ is denoted by ESQM$_{\rm ls}$ with the $\delta$ specified, and recall that SPGL1 is the algorithm used for obtaining initial points for FPA$_{\rm retract}$ and ESQM$_{\rm ls}$ by solving the auxiliary problem~\eqref{SPGL1nonconvexL1}. As before, we present the time for performing QR decomposition on $A^T$ (denoted by QR), the time for computing $A^\dagger b$ given the QR factorization of $A^T$ (denoted by Slater), the CPU time of algorithms (CPU time),\footnote{The reported CPU times of FPA$_{\rm retract}$ and ESQM$_{\rm ls}$ do not include the time for generating $\xfeasss$ and $x^0$.} the number of iterations (denoted by Iter), the recovery errors $\text{RecErr} := \frac{\|x^* - \xorig\|}{\max\{1, \|\xorig\|\}}$ and the residuals ${\rm Residual} := \frac{\|Ax^* - b\|_{LL_2,\gamma} - \overline{\sigma} }{\overline{\sigma}}$, where $x^*$ is the approximate solution returned by the respective algorithm.

>>>>

\section{Introduction}
We consider the following difference-of-convex (DC) optimization problem with smooth inequality and geometric constraints:
\begin{equation}\label{P0}
  \begin{array}{rl}
\min\limits_{x\in\R^n} & P(x) := P_1(x) - P_2(x)\\
{\rm s.t.} & g_i(x)\leq 0,~~ i = 1,\ldots, m,\\
           & x\in C,
  \end{array}
\end{equation}
where $P_1:\R^n\rightarrow\R$ and $P_2:\R^n\rightarrow\R$ are {\color{blue}finite-valued convex} functions, $C\subseteq \R^n$ is a nonempty compact convex set, $g_i:\R^n\rightarrow\R$ is smooth with Lipschitz gradient on $C$ for each $i$, and the feasible set $\mathcal{F} :=\{x\in C:\; g_i(x)\le 0, ~ i = 1,\ldots, m\}$ is nonempty. This model problem covers many important optimization problems arising from diverse areas. For example, one can model the compressed sensing (CS) problem with Cauchy noise by minimizing the difference of $\ell_1$-norm and $\ell_2$-norm \cite{YiLH15} or its variants subject to an inequality constraint defined by the so-called Lorentzian norm \cite{CaBA10}, and the set $C$ in \eqref{P0} can be used for modeling other prior information such as bounds or nonnegativity; more discussions on this particular application can be found in Section~\ref{sec6.2} below. Problem~\eqref{P0} is typically challenging to solve because it is in general nonconvex and nonsmooth.% optimization problem.

@@

More explicitly, we consider the following special case of \eqref{P0}:
\begin{equation}\label{P1}
  \begin{array}{rl}
\min\limits_{x\in\R^n} & P(x) := P_1(x) - P_2(x)\\
{\rm s.t.} & \ell(Ax - b)\leq \sigma,\\
           & x\in C,
  \end{array}
\end{equation}
where $P_1$ and $P_2$ are {\color{blue}finite-valued convex} functions, $C$ is a nonempty compact convex set, $A\in\R^{p\times n}$, $b\in\R^p$,  $\sigma >0$, the set ${\cal F}_1:= \{x\in C:\; \ell(Ax - b)\le \sigma\}\neq \emptyset$, and $\ell$ also satisfies the following Assumption~\ref{gassum}:
\begin{assumption}\label{gassum}
The function $\ell:\R^p\to \R$ is defined as
\[
\ell(u):= \sum_{i=1}^{p} {\color{blue} \varphi}(u_i^2),
\]
where ${\color{blue} \varphi}:\R_+\rightarrow \R_+$ satisfies the following conditions:
\begin{enumerate}[{\rm (i)}]
  \item ${\color{blue} \varphi}$ is a continuous concave function with ${\color{blue} \varphi}(0) = 0$;
  \item ${\color{blue} \varphi}$ is continuously differentiable with positive derivative on $(0, \infty)$;
  \item The limit $\vartheta:= \lim\limits_{t\downarrow 0}{\color{blue} \varphi}'(t)$ exists, and the function $t\mapsto {\color{blue} \varphi}'_+(t)$ is locally Lipschitz on $[0,\infty)$.\footnote{Here, and throughout this paper, we use ${\color{blue} \varphi}'_+(t)$ to denote the right hand derivative of ${\color{blue} \varphi}$ at $t$, which is defined as ${\color{blue} \varphi}'_+(t)=\lim\limits_{h \rightarrow 0^+}\frac{{\color{blue} \varphi}(t+h)-{\color{blue} \varphi}(t)}{h}$.}
\end{enumerate}
\end{assumption}

@@

{\color{blue}We compare Algorithm~\ref{SQP_retract2} and Algorithm~\ref{ESQM} on solving random instances of \eqref{E4}. Specifically, we generate two matrices ${\bf A}_{\rm re}\in \R^{p\times n}$ and ${\bf A}_{\rm im}\in \R^{p\times n}$ with i.i.d standard Gaussian entries, let
\begin{equation*}
A := \left[ \begin{matrix}
{\bf A}_{\rm re} & -{\bf A}_{\rm im}\\
{\bf A}_{\rm im} & {\bf A}_{\rm re}
\end{matrix} \right]
\end{equation*}
and then normalize each column of $A$. Next, we generate two vectors ${\bf u}\in\R^k$ and ${\bf v}\in\R^k$ with i.i.d standard Gaussian entries. We choose an index set $I$ of size $k$ at random and define the original sparse signal ${\bf z}_{\rm orig}\in {\mathbb{C}}^{n}$ by setting ${\bf z}_j := {\bf u}_j + i {\bf v}_j$ for all $j \in I$ and ${\bf z}_j = 0$ otherwise. We then let
\[
\xorig = \begin{bmatrix}
  {\bf z}_{\rm re}\\
  {\bf z}_{\rm im}
\end{bmatrix}\ \ {\rm and}\ \ b = A\xorig + 0.005\hat{\epsilon},
\]
where ${\bf z}_{\rm re}$ and ${\bf z}_{\rm im}$ are respectively the real and imaginary parts of ${\bf z}_{\rm orig}$, and $\hat{\epsilon}$ is a random vector with each entry independently following the Cauchy$(0, 1)$ distribution, i.e., $\hat{\epsilon}_{j} = \tan(\pi(\widetilde{\epsilon}_j - \frac{1}{2}))$ with each $\widetilde{\epsilon}_j\in\R^{2p}$ being uniformly chosen in $[0, 1]$. Finally, we set $\overline{\sigma} = 1.2\|0.005\hat{\epsilon}\|_{LL_2,\gamma}$ with $\gamma = 0.05$.}

@@

In our numerical experiments, we let $\mu =0.95$, $j = 2$, $\gamma = 0.05$, {\color{blue} $(p,n,k) = (360i,1280i,60i)$ }with $i\in \{2, 4, 6, 8, 10\}$. For each $i$, we generate {\color{blue}20 random instances} as described above. The computational results averaged over the {\color{blue} $20$ random instances} are shown in Table~\ref{table2}. In the table, Algorithm~\ref{SQP_retract2} is denoted by FPA$_{\rm retract}$, Algorithm~\ref{ESQM} with different $\delta$ is denoted by ESQM$_{\rm ls}$ with the $\delta$ specified, and recall that SPGL1 is the algorithm used for obtaining initial points for FPA$_{\rm retract}$ and ESQM$_{\rm ls}$ by solving the auxiliary problem~\eqref{SPGL1nonconvexL1}. As before, we present the time for performing QR decomposition on $A^T$ (denoted by QR), the time for computing $A^\dagger b$ given the QR factorization of $A^T$ (denoted by Slater), the CPU time of algorithms (CPU time),\footnote{The reported CPU times of FPA$_{\rm retract}$ and ESQM$_{\rm ls}$ do not include the time for generating $\xfeasss$ and $x^0$.} the number of iterations (denoted by Iter), the recovery errors $\text{RecErr} := \frac{\|x^* - \xorig\|}{\max\{1, \|\xorig\|\}}$ and the residuals ${\rm Residual} := \frac{\|Ax^* - b\|_{LL_2,\gamma} - \overline{\sigma} }{\overline{\sigma}}$, where $x^*$ is the approximate solution returned by the respective algorithm.
